---
title: "Visualizing Airbnb Listings in San Francisco"
output:
  html_document:
    df_print: paged
---
**Task:** Exploratory Analysis Using ggplot2

**Dataset:** Airbnb listings in San Francisco, the dtaset contains 16 variables and 6998 observations. The data can be obtained from [http://data.insideairbnb.com/united-states/ca/san-francisco/2020-10-05/visualisations/listings.csv]


load required libraries
```{r}
library(tidyverse)

```
**Data Preparation and Preprocessing**

Import data and check on its structure
```{r}
listings <- read.csv("listings.csv")
str(listings)
```

Change data types
```{r}
listings$last_review <- as.Date(listings$last_review) #change data type of last_review to Date
listings$room_type <- factor(listings$room_type) #room_type data type from character to factor
listings$neighbourhood <- factor(listings$neighbourhood) #neighbourhood data type from character to factor
listings$host_id <- factor(listings$host_id) #host_id from numeric to factor

```
Drop redundant columns and get the summary statistics
```{r}
listings <- listings[-c(1,4)] #drop id and host_name columns, id is insignificant and host_name for ethical reasons
summary(listings) # get the summary of the data
```
Handling missing values, 
```{r}
listings$neighbourhood_group <- NULL # drop neighbourhood_group column because it has all NA"s
listings$last_review <- NULL #drop last_review column, there are lots of NA's because if there were no reviews for the listings, the date doesn't exist
listings$reviews_per_month[is.na (listings$reviews_per_month)] <- 0 #replace the NA's in reviews_per_month with zero since there were no reviews for that particular day

```

**Exploratory Data Analysis:** VIsualization using ggplot2

```{r Room_type}
#distribution of room_type
ggplot(listings, aes(x=room_type, fill=room_type)) + geom_bar() + labs(title = "Room_type")
```
Entire home/apartment is the most preferred by tourists followed by Private room. the least preferred is the Hotel room

```{r neighbourhood}
#top 10 frequent neighbourhoods
listings %>% count(neighbourhood) %>% arrange(desc(n))%>% head(10) %>% ggplot(aes(x=neighbourhood, y=n, fill=neighbourhood)) +geom_bar(stat = "identity",position = "dodge") + coord_flip() + ylab("frequency") + labs(title = "Top 10 neighbourhoods")
```

From the bar chart, it can be deduced that, Donwntown/Civic Centre is the most preferred neighbourhood by tourists

```{r price}
#variation in Price
ggplot(listings, aes(x=price)) + geom_histogram(fill="aquamarine", bins = 100, color= "black") + labs(title = "Price") #the histogram is not that good

df=data.frame(price=listings["price"][listings["price"]<=2000])
ggplot(df, aes(x= price)) + geom_histogram(fill="aquamarine", bins = 100, color="black") + labs(title = "Price <= 2000") #plot price <= 2000 for some clarity

```
THe first plot is not clear because of outliers
the second plot is a bit clear where I considered the price of <=2000
```{r minimum_nights}
df2= data.frame(minimum_nights=listings["minimum_nights"][listings["minimum_nights"]<=30])
ggplot(df2, aes(x=minimum_nights)) + geom_histogram(bins=50, fill="orange", color= "black")
```
minimum night of 30 has the highest listing

```{r price vs minimum_nights}
listings %>% mutate(minimum_nights = ifelse(minimum_nights<=30, "short_term_rentals", "long_term_rentals")) %>% ggplot(aes(x=minimum_nights, y=price, fill= minimum_nights)) + geom_bar(stat = "identity")
```
Short term rentals of <=30 nights are more expensive than long term rentals maybe because they have high listings

```{r}
#hosts running business with multiple listings
listings %>% group_by(neighbourhood) %>% count(host_id) %>% arrange(desc(n))%>% head(10) %>% ggplot(aes(x=host_id, y=n, fill=neighbourhood)) +geom_bar(stat = "identity", position="dodge") + coord_flip() + labs(title="Hosts running business with multiple listings")
```
The host with id 4430421 has the most listings and in 2 neighbourhoods

```{r}
#distibution of room_type across the region
ggplot(listings,aes(x=longitude, y= latitude, color=room_type)) + geom_point()
```

